% Latex-template for one-page abstract submissions.
% NUMTA2026 - Numerical Computations: Theory and Algorithms
% 21-27 June 2026, Falkensteiner Garden Calabria Resort, Curinga (Italy)

\documentclass[oribibl]{llncs}

% Please, do not use your own macros/redefinitions.

%\usepackage[russian]{babel}

\usepackage{fancyhdr}
\renewcommand{\headrulewidth}{0pt}
\renewcommand{\footrulewidth}{0pt}

\begin{document}
\cleardoublepage

\title{Finding a Zero Value Point of a Non-negative Univariate Function}
\author{Konstantin Barkalov, Alexander 	, Roman Strongin}

%%% Please note that the first author is supposed to be the speaker.

\institute{Lobachevsky State University of Nizhny Novgorod, Nizhny Novgorod, Russia \\
\email{barkalov@vmk.unn.ru, a.zaytsev@unn.ru, strongin@unn.ru}}

\maketitle

\thispagestyle{fancy}

\textbf{Keywords.} Global optimization; multiextremal functions; Lipschitz condition.

 \vspace*{0.5cm}

%-----------------------------------------------------------------

This work considers one-dimensional Lipschitz global optimization problems and methods for their solution. The importance of these problems is due to the following reasons.
There are a significant number of practical applications where it is necessary to solve univariate global optimization problems. It is also important to develop  one-dimensional methods since they can be successfully generalized to multidimensional case using various schemes \cite{Sergeyev2013,Sergeyev2017}.

In general, the assumption of Lipschitz continuity of the objective function is typical for many approaches in developing algorithms for global optimization (see, e.g., \cite{Evtushenko2013,Jones2021}). Proper functioning of such algorithms (i.e., convergence to the global minimum) is ensured either by using a sufficiently good estimate of the Lipschitz constant or by considering a set of possible values of this constant.

The present paper proposes a new algorithm designed to solve global optimization problems where the value of the objective function at its minimum point is assumed known but the location of that point needs to be determined. While well-known Lipschitz optimization methods could certainly be applied to solve such problems, unlike them, the proposed new algorithm guarantees convergence to the global minimum of the objective function without requiring any estimation of the Lipschitz constant's value.

\textbf{Acknowledgements.}
This work was supported by the Ministry of Science and Education of the Russian Federation, project no. 075-02-2026-1346.

 \vspace{0.5cm}

\begin{thebibliography}{4}

%\bibitem{Book} Cantor G. (1955) \emph{Contributions to the founding of the theory of transfinite numbers}. Dover Publications, New York.
\bibitem{Sergeyev2013} Sergeyev Y.D., Strongin R.G., Lera D. (2013) \emph{Introduction to global optimization exploiting space-filling curves}. Springer, New York.

\bibitem{Sergeyev2017} Sergeyev Y.D., Kvasov D.E. (2017) \emph{Deterministic global optimization: An introduction to the diagonal approach}. Springer, New York.

%\bibitem{Article} Gordon P. (2004) Numerical cognition without words: {E}vidence from {A}mazonia. \emph{Science}, Vol.~306, pp.~496--499.

\bibitem{Evtushenko2013} Evtushenko Yu.G., Posypkin M.A. (2013) A deterministic approach to global box-constrained optimization. \emph{Optimization Letters}, Vol.~7(4), pp.~819--829. 

\bibitem{Jones2021} Jones D.R., Martins J.R.R.A. (2021) The DIRECT algorithm: 25 years Later. \emph{J. Glob. Optim.}, Vol.~79. pp.~521--566.

%\bibitem{Proceedings} Author I., Author I.I. (2019) An abstract. In Proceedings of the \emph{International Conference``XXX''} (ed. by Editor E.), Place (Country), p.~1.

%\bibitem{BookChapter} Author X. (2019) A chapter. In \emph{YYY Book} (ed. by Editor E.), Place (Country), pp.~10--20.


\end{thebibliography}

\end{document}
