\documentclass[a4paper]{locconf}
\usepackage{graphicx}
\usepackage[russian,english]{babel}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amssymb}
\begin{document}
\title{High Performance Computing for Global Optimization Problems}


\author{K.A. Barkalov, V.P. Gergel}

\address{Lobachevsky State University of Nizhni Novgorod, Gagarin avenue 23, Nizhni Novgorod, Russia, 603950}



\begin{abstract}
All articles must contain an abstract. The abstract should give readers concise information about the content of the article and indicate the main results obtained and conclusions drawn. The abstract is not part of the text and should be complete in itself; no table numbers, figure numbers, references or displayed mathematical expressions should be included. It should be suitable for direct inclusion in abstracting services and should not normally exceed 200 words in a single paragraph. Since contemporary information-retrieval systems rely heavily on the content of titles and abstracts to identify relevant articles in literature searches, great care should be taken in constructing both.
\end{abstract}

\section{Introduction}

This file should be used as a pattern for English Thesis. You can write on email mingazov88@gmail.com with questions concerning the preparation of the English Thesis.

\section{Problem statement}

Let us consider the $N$-dimensional global optimization problem
\begin{equation}\label{problem}
\varphi(y^\ast)=\min{\left\{\varphi(y): y \in D, \; g_i(y)\leq 0, \; 1 \leq i \leq m\right\}}
\end{equation}
with search domain
\begin{equation}\label{D}
D=\left\{y\in R^N: a_j\leq y_j \leq b_j, 1\leq j \leq N \right\}.
%D=\left\{y\in R^N: a_i\leq y_i \leq b_i, 1\leq i \leq N\right\}.
\end{equation}

The objective function $\varphi(y)$ (henceforth denoted by $g_{m+1}(y)$) and
the left-hand sides $g_i(y), \; 1\leq i \leq m$, of the constraints
satisfy the Lipschitz conditions with constants $L_i, \; 1 \leq i \leq
m+1$, respectively, and may be multiextremal. It is assumed that the
functions $g_i(y),\; 1 \leq i \leq m+1$, are defined and computable only in the corresponding domains
\[
Q_1=D, \; Q_{i+1}=\left\{y \in Q_i : g_i(y) \leq 0 \right\}, \; 1 \leq i \leq m.
\]

These conditions allow for the introduction of a classification of the
points $y \in D$ according to the number $\nu (y)$ of the constraints
computed at this point. 
%отсюда

Thus, a \textit{trial} at a point $y^k \in D$ executed at the $k$-th
iteration of the algorithm will consist in computing the values $g_1(y),...,g_\nu(y),$ where the
index $\nu \leq m$ is determined by the conditions
	\[
		g_i(y^k )\leq 0, \; 1 \leq i < \nu, \; g_\nu(y^k)>0, \; \nu \leq m.
	\]
The occurrence of the first violation of the constraint terminates the
trial. In the case when the point $y^k$ is a feasible one, i.e. when
$y \in Q_{m+1}$, the trial includes the computation of the values of
all the functions of the problems and the index is assumed to be
$\nu=m+1$. The pair of values
\[
\nu=\nu(y^k), \; z^k=g_\nu(y^k)
\]
is a \textit{trial result}.

\Russian
Параллельный индексный алгоритм, который может быть применен для решения подобных задач с частично вычислимыми функциями и который реализован в системе Globalizer, подробно описан в [2]. Основная идея алгоритма состоит в сведении исходной многомерной задачи к набору связанных подзадач меньшей размерности, и параллельному их решению. Схемы редукции размерности, используемые при работе индексного алгоритма, кратко описаны в следующем разделе.



\section*{Acknowledgments}
This study was supported by the Russian Science Foundation, project No 16-11-10150.

\section*{References}

\medskip

\begin{thebibliography}{9}

\bibitem{Strongin2000}
Strongin, R.G. Global optimization with non-convex constraints. Sequential and parallel algorithms. / R.G. Strongin, Ya.D.  Sergeyev --- Dordrecht: Kluwer Academic Publishers, 2000. --- 704 p.

\bibitem{Strongin2013}
Strongin, R.G. Parallel computing in global optimization problems / R.G. Strongin, V.P. Gergel, V.A. Grishagin, K.A. Barkalov --- Moscow: Publishing of the Moscow State University, 2013. --- 280 p. --- (in Russian).

\bibitem{Jones}
Jones, D.R. The DIRECT global optimization algorithm / In: Floudas, C. A., Pardalos, P. M. (eds.) The Encyclopedia of Optimization, Second Edition. --- Heidelberg: Springer, 2009. --- P. 725--735. 

\bibitem{Zilinskas2011}
Paulavi\v{c}ius, R. Parallel branch and bound for global optimization with combination of Lipschitz bounds / R. Paulavi\v{c}ius, J. \v{Z}ilinskas, A. Grothey // Optimization Methods \& Software. --- 2011. --- Vol. 26(3). --- P. 487--498.

\bibitem{Evtushenko}
Evtushenko, Yu.G. Parallel global optimization of functions of several variables / Yu.G. Evtushenko, V.U. Malkova, A.A. Stanevichyus // Computational Mathematics and Mathematical Physics. --- 2009. --- Vol. 49 (2). --- P. 246--260.

\bibitem{Sysoyev}
Sysoyev, A.V. MPI implementation of dimension reduction multilevel scheme for parallel solving the global optimization problems / A.V. Sysoyev, K.A. Barkalov, V.P. Gergel, I.G. Lebedev // Russian Supercomputing Days: Proceedings of the International Scientific Conference. --- Moscow: Publishing of the Moscow State University, 2015. --- P. 61--68.

\bibitem{BarkalovLebedev}
Barkalov, K.A. Solving global optimization problems on GPU / K.A. Barkalov, I.G. Lebedev // Russian Supercomputing Days: Proceedings of the international conference. --- Moscow: Publishing of the Moscow State University, 2016. --- P. 640--650.

\bibitem{BarkalovGergel}
Barkalov, K. Parallel global optimization on GPU / K. Barkalov K., V. Gergel // Journal of Global Optimization. --- 2016. --- Vol. 66(1). --- P. 3--20.

\bibitem{BarkalovGergelLebedev}
Barkalov, K. Use of Xeon Phi Coprocessor for Solving Global Optimization Problems / K. Barkalov, V. Gergel, I. Lebedev //  Lecture Notes in Computer Science. --- 2015. --- Vol. 9251. --- P. 307--318.

\bibitem{Gergel2017}
Gergel, V. An Approach for Generating Test Problems of Constrained Global Optimization / V. Gergel // Lecture Notes in Computer Science. --- 2017. --- Vol. 10556. --- P. 314--319.

\bibitem{Gergel2016}
Gergel, V. Adaptive nested optimization scheme for multidimensional global search / V. Gergel, V. Grishagin, A. Gergel // Journal of Global Optimization. --- 2015. --- Vol. 66(1). --- P. 35--51.

\bibitem{Sergeyev2017}
Sergeyev, Ya.D. Deterministic global optimization: An introduction to the diagonal approach / Ya.D. Sergeyev, D.E. Kvasov --- New York: Springer, 2017. --- 136 p.

\end{thebibliography}

\end{document}


